---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# simkit

<!-- badges: start -->
[![R-CMD-check](https://github.com/sohamsaha99/simkit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sohamsaha99/simkit/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

simkit provides a tidy, pipeline-based framework for statistical simulation
studies. It handles parallelism, checkpointed storage, fault tolerance,
progress reporting, and performance evaluation so you can focus on
estimators and scenarios.

## Installation

You can install the development version of simkit from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("sohamsaha99/simkit")
```

## Example

A `SimDesign` is assembled by adding scenarios, methods, and configuration
together with `+`, then run and evaluated:

```{r example}
library(simkit)
library(future)
plan(sequential)

design <- sim_design(replicates = 20, seed = 1) +
  sim_scenario(
    name   = "n30",
    params = list(n = 30, mu = 0),
    dgp    = function(params) rnorm(params$n, params$mu),
    truth  = function(params) list(mean = params$mu)
  ) +
  sim_method(name = "sample_mean", fn = function(data, params) list(est = mean(data))) +
  sim_storage(path = tempfile("simkit_readme_"))

design |>
  sim_run(progress = FALSE) |>
  sim_collect() |>
  sim_evaluate(bias = function(sims, truth) mean(sims$est) - truth$mean)
```

See `vignette("getting-started")` for the full pipeline, including scenario
grids, bootstrap variance estimation, and plotting.
