---
title: "Getting Started with simkit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with simkit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(simkit)
library(future)
library(ggplot2)

plan(sequential)
```

`simkit` structures a Monte Carlo simulation study as a pipeline: describe
*what* to simulate and *how* to estimate it, run it once, then evaluate and
plot the results. This vignette walks through that pipeline end to end,
comparing three estimators of the mean under normal and heavy-tailed errors.

To keep this vignette fast to build, it uses `replicates = 500` and a small
bootstrap size. In a real study you would typically use several hundred to a
few thousand replicates.

## 1. Design

A `SimDesign` is assembled by adding components together with `+`. Start with
`sim_design()`, then add a scenario grid, one or more methods, and a storage
location.

`sim_scenario_grid()` expands a list of parameters into one named scenario
per combination:

```{r design}
storage_path <- tempfile("simkit_getting_started_")

design <- sim_design(
    replicates  = 500,
    seed        = 42,
    label       = "mean_estimators_v1",
    description = "Comparing sample mean, trimmed mean, and median under
                   normal and heavy-tailed errors."
  ) +

  sim_scenario_grid(
    params  = list(n = c(30, 100), dist = c("normal", "t3")),
    name_fn = function(p) paste0(p$dist, "_n", p$n),
    dgp     = function(p) {
      if (p$dist == "normal") rnorm(p$n, mean = 2, sd = 1)
      else                    2 + rt(p$n, df = 3)
    },
    truth = function(p) list(mean = 2)
  ) +

  sim_method(
    name = "sample_mean",
    fn   = function(data, params) {
      se <- sd(data) / sqrt(params$n)
      list(est = mean(data), se = se,
           ci_lo = mean(data) - 1.96 * se,
           ci_hi = mean(data) + 1.96 * se)
    }
  ) +

  sim_method(
    name = "trimmed_mean",
    fn   = sim_bootstrap(
      estimator = function(data, params) list(est = mean(data, trim = 0.1)),
      B = 50, statistic = "est", output = c("se", "ci")
    )
  ) +

  sim_method(
    name = "median",
    fn   = sim_bootstrap(
      estimator = function(data, params) list(est = median(data)),
      B = 50, statistic = "est", output = c("se", "ci")
    )
  ) +

  sim_storage(path = storage_path, resume = TRUE)

design
```

## 2. Validate

Before committing to a full run, `sim_check()` runs a handful of trial
replicates for every (scenario, method) pair in an ephemeral scratch
directory -- never the design's configured storage path -- and reports
pass/fail with timing:

```{r check}
sim_check(design, replicates = 2)
```

## 3. Run

`sim_run()` executes every replicate, writing one `.rds` file per
(scenario, method, replicate) to the storage path as it goes. A deliberate
per-replicate error is recorded as a sentinel file rather than halting the
run (the default `on_error = "warn"`), and rerunning with
`sim_storage(resume = TRUE)` skips replicates already on disk.

```{r run}
results <- sim_run(design, progress = FALSE)
results
```

## 4. Evaluate

`sim_collect()` reads the stored replicates back into a single tidy tibble,
and `sim_evaluate()` applies one or more user-supplied metric functions to
every (scenario, method) pair, comparing simulated output against the
scenario's `truth`:

```{r evaluate}
evaluation <- results |>
  sim_collect() |>
  sim_evaluate(
    bias     = function(sims, truth) mean(sims$est) - truth$mean,
    mse      = function(sims, truth) mean((sims$est - truth$mean)^2),
    variance = function(sims, truth) var(sims$est),
    coverage = function(sims, truth) mean(sims$ci_lo <= truth$mean &
                                          truth$mean <= sims$ci_hi)
  )

evaluation
```

`sim_summarize()` pivots the long-format evaluation into one row per
(scenario, method) with a column per metric:

```{r summarize}
sim_summarize(evaluation)
```

## 5. Report

`sim_plot()` extracts scenario parameters from their names and maps them to
plot aesthetics, returning an ordinary `ggplot` object that can be extended
with further layers:

```{r plot, fig.width = 6, fig.height = 4}
evaluation |>
  sim_plot(metric = "mse", x = "n", color = "method") +
  labs(title = "MSE of mean estimators under normal vs. heavy-tailed errors")
```

## 6. Revisit later

`sim_catalog()` indexes every experiment found under a root directory, and
`sim_load()` recovers a `SimDesign` and its results from a catalog entry --
handy for coming back to a study in a fresh R session without re-specifying
the design:

```{r catalog}
catalog_root <- dirname(storage_path)
sim_catalog(catalog_root)

sim_catalog(catalog_root) |>
  sim_load() |>
  sim_collect() |>
  head()
```

```{r cleanup, include = FALSE}
unlink(storage_path, recursive = TRUE)
```
